---
title: "AKQoL Rasch analysis"
subtitle: "Actinic Keratosis Quality of Life"
title-block-banner: "#009ca6"
title-block-banner-color: "#FFFFFF"
author: 
  name: 'Magnus Johansson'
  affiliation: 'RISE Research Institutes of Sweden'
  affiliation-url: 'https://www.ri.se/kbm'
  orcid: '0000-0003-1669-592X'
date: last-modified
date-format: YYYY-MM-DD
format:
  html:
    theme: materia
    toc: true
    mainfont: 'Lato'
    monofont: 'Roboto Mono'
    code-overflow: wrap
    code-tools: true
    code-fold: true
    code-link: true
    number-sections: true
    fig-dpi: 96
    layout-align: left
    linestretch: 1.6
    license: CC BY
execute: 
  cache: true
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

## Itemlabels

This is actually created primarily using copy&paste from the Word-file available in "Supporting Information" in the paper. I changed the variable names and the first variable to be character format rather than numeric.

```{r}
#install.packages("datapasta")
# use the "paste as dataframe" in the "Addins" meny above, after installing the library "datapasta"
itemlabels <- data.frame(
  stringsAsFactors = FALSE,
                itemnr = paste0("q",c(1:9)),
                item = c("During the last week I have been annoyed with having to protect my skin each time I’m going out in the sun.",
                       "During the last week my sun damaged skin has given me a reason to think about what is important in life.",
                       "During the last week my quality of life has been decreased because of my sun damaged skin.",
                       "During the last week I have feared that the sun damage of my skin will develop into a more serious skin disease.",
                       "During the last week I have tried to hide my sun damaged skin from other people with e.g. make-up or clothes.",
                       "During the last week I have felt guilt in relation to my sun damaged skin.",
                       "During the last week I have looked at my skin and examined it for sun damages.",
                       "During the last week my life has been more difficult because of the sun damages of my skin.",
                       "During the last week I have thought about how to behave in the sun.")
)
```

## Import data

```{r}
### If needed, install the devtools package:
# install.packages('devtools')
### Then install the RISEkbmRasch package and its dependencies:
# devtools::install_github("pgmj/RISEkbmRasch", dependencies = TRUE)
library(RISEkbmRasch)
library(skimr)
library(TAM)
library(psych)

sc <- read.csv("scqol.csv")

scItems <- sc %>% 
  select(starts_with("SCQ")) %>% 
  select(!ends_with("Q10"))

dif.sex <- factor(sc$Sex)
dif.age <- sc$Age

names(scItems) <- paste0("q",c(1:9))
```

## SCQoL

### Descriptives

::: {.panel-tabset}
#### Demographics
```{r}
RIdemographics(dif.sex,"Sex")
```

```{r}
hist(dif.age, main = "Age", col = "lightblue", xlab = "Age in years")

describe(dif.age) %>% 
  select(n,mean,median,min,max,range,skew,kurtosis) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  kbl_rise()
```
#### Ordinal sum scores
```{r}
RIrawdist(scItems)
```

#### Missing/item
```{r}
RImissing(scItems)
```
#### Missing/person
```{r}
RImissingP(scItems)
```
#### Tile plot
```{r}
RItileplot(scItems)
```
#### Stacked bars
```{r}
RIbarstack(scItems)
```
#### Guttman heatmap
```{r}
RIheatmap(scItems)
```
#### Barplots
```{r}
#| layout-ncol: 2
RIbarplot(scItems)
```
#### Expected Value Curves
```{r}
#| layout-ncol: 2
tam1 <- tam(as.matrix(scItems), irtmodel = "PCM", verbose = FALSE)
plot(tam1)
```
#### Skim
```{r}
skim(scItems)
```
#### All responses
```{r}
RIallresp(scItems)
```
:::

### Rasch analysis 1

::: {.panel-tabset}
#### Item fit
```{r}
RIitemfitPCM(scItems)
```
#### PCA of residuals
```{r}
RIpcmPCA(scItems)
```
#### Loadings on 1st contrast
```{r}
RIloadLoc(na.omit(scItems))
```
#### Residual correlations
```{r}
RIresidcorr(scItems, 0.2)
```
#### Response categories
```{r}
#| layout-ncol: 2
RIitemCats(scItems)
```
#### Targeting
```{r}
RItargeting(scItems)
```
#### Item hierarchy
```{r}
#| fig-height: 8
RIitemHierarchy(scItems)
```
:::

### DIF

#### Gender

```{r}
RIdifTable(scItems, dif.sex)
```

#### Age
```{r}
RIdifTable(scItems, dif.age)
```


## AKQoL

```{r}
ak <- read.csv("akqol.csv")

ak.sex <- factor(ak$Sex)

akItems <- ak %>% 
  select(starts_with("AKQ"))

names(akItems) <- paste0("q",c(1:9))

```

### Descriptives

::: {.panel-tabset}
#### Tile plot
```{r}
RItileplot(akItems)
```
#### Stacked bars
```{r}
RIbarstack(akItems)
```
#### Guttman heatmap
```{r}
RIheatmap(akItems)
```
#### Barplots
```{r}
#| layout-ncol: 2
RIbarplot(akItems)
```
#### Expected Value Curves
```{r}
#| layout-ncol: 2
tam1 <- tam(as.matrix(akItems), irtmodel = "PCM", verbose = FALSE)
plot(tam1)
```
#### Skim
```{r}
skim(akItems)
```
#### All responses
```{r}
RIallresp(akItems)
```
:::

### Rasch analysis 1

::: {.panel-tabset}
#### Item fit
```{r}
RIitemfitPCM(akItems)
```
#### PCA of residuals
```{r}
RIpcmPCA(akItems)
```
#### Loadings on 1st contrast
```{r}
RIloadLoc(na.omit(akItems))
```
#### Residual correlations
```{r}
RIresidcorr(akItems, 0.2)
```
#### Response categories
```{r}
#| layout-ncol: 2
RIitemCats(akItems)
```
#### Targeting
```{r}
RItargeting(akItems)
```
#### Item hierarchy
```{r}
#| fig-height: 8
RIitemHierarchy(akItems)
```
:::

### DIF

#### Gender

```{r}
RIdifTable(akItems, ak.sex)
```
